---
output:
  pdf_document: default
---

# Nate Silver 538 Masculinity Survey
538.com published an article on masculinty based on a survey they took with WNYC Studios.
<br>Since most of the article focuses on visualization (and they were kind enough to make the data available), I wanted to take a look and see what trends might come out through ML-based analytics.

https://fivethirtyeight.com/features/what-do-men-think-it-means-to-be-a-man/

### The workflow I used can be broken up into 3 sections:
### 1. Data wrangling
### 2. Unsupervised analysis using kmeans
### 3. Supervised analysis: What variables drive an idea (or lack of) masculinity?

# Section 1: Setup and Data Import
```{r results=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

library(tidyverse)
library(rlang)
library(caret)
library(recipes)
library(ggfortify)
library(ranger)
```

#### Load data
https://data.fivethirtyeight.com/
```{r message=FALSE, warning=FALSE}
input_1 = read_csv("C:/Users/db1/Documents/R/538_man_survey/raw-responses.csv") %>% 
  select(-1, -StartDate, -EndDate)
data_1 = input_1
```

#### Review unique responses on questions
#### Get some idea of where to start cleanup
```{r}
data_1 %>% 
  select(2) %>% 
  group_by_at(names(.)) %>% 
  summarise(n = n())
```

#### Change "No answer" to NA across dataframe
```{r}
data_1 = data_1 %>% 
  mutate_all(function(x){
    ifelse(x == "No answer", NA, x)
  })

data_1 %>% 
  select(2) %>% 
  group_by_at(names(.)) %>% 
  summarise(n = n())
```

#### Change "Not selected" to 0 in one-hot columns
```{r}
which_dummy = map_lgl(data_1 %>% 
             colnames(),
           function(i){
            data_1 %>%
              select(i) %>% 
               paste() %>% 
              str_detect("Not selected")
          }) %>% 
  which()

data_dummy = data_1 %>% 
  select(which_dummy) %>% 
  mutate_all(function(x){
    ifelse(x == "Not selected", 0, 1)
  })


data_1 = data_1 %>% 
  select(-which_dummy) %>% 
  bind_cols(data_dummy) %>% 
  select(names(data_1))

data_1 %>% 
  select(1:5) %>% 
  slice(1:5)
```

#### Encode ordinal columns
#### Function to encode ordinal columns
```{r}
convert_ordinal_string_column = function(input_data_frame,
                                         column_detection_string,
                                         LHS_strings = NULL, RHS = NULL){
which_ordinal = map_lgl(input_data_frame %>% 
             colnames(),
           function(i){
            data_1 %>%
              select(i) %>% 
               paste() %>% 
              str_detect(column_detection_string)
          }) %>% 
  which()

data_ordinal = input_data_frame %>%
    select(which_ordinal)
  for (i in 1:length(LHS_strings)) {
    data_ordinal = data_ordinal %>% 
    mutate_all(function(x){
      ifelse(str_detect(x, LHS_strings[i]), RHS[i], x)
    }
    )
  }
data_ordinal = data_ordinal %>%
    mutate_all(function(x){
      ifelse(x %in% RHS, x, NA)
    })

input_data_frame = input_data_frame %>% 
  select(-which_ordinal) %>% 
  bind_cols(data_ordinal) %>% 
  select(names(input_data_frame))

return(input_data_frame)

}
```
#### Call function to encode ordinal columns on individual questions
```{r}
# q0001, q0002
data_1 = data_1 %>% 
  convert_ordinal_string_column("Somewhat",
                                LHS_strings = c("Very", 
                                                "Somewhat", 
                                                "Not at all",
                                                "Not very",
                                                "Not too"), 
                                RHS = c(4, 3, 1, 2, 2))

data_1 = data_1 %>% 
  convert_ordinal_string_column("\\$0",
                                LHS_strings = c("\\$0",
                                                "\\$10,",
                                                "\\$25,",
                                                "\\$50,",
                                                "\\$75,",
                                                "\\$100,",
                                                "\\$125,",
                                                "\\$150,",
                                                "\\$175,",
                                                "\\$200,",
                                                "Prefer "), 
                                RHS = c(5, 17.5, 37.5, 62.5, 87.5, 112.5, 
                                        137.5, 162.5, 187.5, 225, NA))

data_1 = data_1 %>% 
  convert_ordinal_string_column("Yes",
                                LHS_strings = c("Yes",
                                                "No"), 
                                RHS = c(1, 0))

# q0007
data_1 = data_1 %>% 
  convert_ordinal_string_column("Never, but open to it",
                                LHS_strings = c("Often",
                                                "Rarely",
                                                "Sometimes",
                                                "Never, but open to it",
                                                "Never, and not open to it"), 
                                RHS = c(5, 4, 3, 2, 1))

# q0009
data_1 = data_1 %>% 
  convert_ordinal_string_column("Not employed, NOT looking for work",
                                LHS_strings = c("Not employed, NOT looking for work",
                                                "Not employed, looking for work",
                                                "Employed, working part-time",
                                                "Not employed, student",
                                                "Not employed-retired",
                                                "Employed, working full-time"), 
                                RHS = c(5, 4, 3, 2, 1, 0))

# q0014
data_1 = data_1 %>% 
  convert_ordinal_string_column("Nothing at all",
                                LHS_strings = c("A lot",
                                                "Some",
                                                "Only a little",
                                                "Nothing at all"), 
                                RHS = c(3, 2, 1, 0))

# q0018
data_1 = data_1 %>% 
  convert_ordinal_string_column("Sometimes",
                                LHS_strings = c("Always",
                                                "Often",
                                                "Sometimes",
                                                "Rarely",
                                                "Never"), 
                                RHS = c(4, 3, 2, 1, 0))

# q0024
data_1 = data_1 %>% 
  convert_ordinal_string_column("Never married",
                                LHS_strings = c("Married",
                                                "Widowed",
                                                "Never married",
                                                "Separated",
                                                "Divorced"), 
                                RHS = c(2, 1, 0, -1, -2))

# q0026
data_1 = data_1 %>% 
  convert_ordinal_string_column("Bisexual",
                                LHS_strings = c("Straight",
                                                "Bisexual",
                                                "Gay",
                                                "Other"), 
                                RHS = c(3, 2, 1, 0))

# q0029
data_1 = data_1 %>% 
  convert_ordinal_string_column("Associate's degree",
                                LHS_strings = c("Post graduate degree",
                                                "College graduate",
                                                "Associate's degree",
                                                "Some college",
                                                "High school or G.E.D.",
                                                "Did not complete high school"), 
                                RHS = c(5, 4, 3, 2, 1, 0))

# race2
data_1 = data_1 %>% 
  convert_ordinal_string_column("Non-white",
                                LHS_strings = c("Non-white",
                                                "White"), 
                                RHS = c(0, 1))

# educ4
data_1 = data_1 %>% 
  convert_ordinal_string_column("Post graduate degree",
                                LHS_strings = c("Post graduate degree",
                                                "College or more",
                                                "Some college",
                                                "High school or less"), 
                                RHS = c(3, 2, 1))

# educ3
data_1 = data_1 %>% 
  convert_ordinal_string_column("High school or less",
                                LHS_strings = c("College or more",
                                                "Some college",
                                                "High school or less"), 
                                RHS = c(3, 2, 1))

# age3
data_1 = data_1 %>% 
  convert_ordinal_string_column("35 - 64",
                                LHS_strings = c("18 - 34",
                                                "35 - 64",
                                                "65 and up"), 
                                RHS = c(26, 49.5, 72))

# kids
data_1 = data_1 %>% 
  convert_ordinal_string_column("Has children",
                                LHS_strings = c("Has children",
                                                "No children"), 
                                RHS = c(1, 0))

data_1 %>% 
  slice(1:5) %>% 
  select(1:5)
```

#### Encode remaining dummy columns
```{r results=FALSE, warning=FALSE}
# Remove space cause one-hot will become column names
data_1 = data_1 %>% 
  mutate_all(function(x){
    str_replace_all(x, " ", "_")
  }) %>% 
  mutate_all(function(x){
    str_replace_all(x, "_/_", "_")
  })

one_hot_cols = c("q0013", "q0028", "q0030", "q0035", "q0036", "racethn4")
data_2 = data_1 %>% 
  select(one_hot_cols)

data_2 = recipe(~ ., data = data_2) %>% 
  step_dummy(!!! one_hot_cols, one_hot = TRUE) %>% 
  prep(training = data_2) %>% 
  bake(new_data = data_2) %>% 
  bind_cols(data_1) %>% 
  select(-one_hot_cols) %>% 
  mutate_all(function(x) as.double(x))
```
```{r}
data_2 %>% 
  slice(1:5) %>% 
  select(1:5)
```

#### Impute NA's using missRanger
```{r}
data_3_impute = data_2 %>%
  missRanger::missRanger(pmm.k = 10, seed = 12345)

data_3_impute %>% 
  slice(1:5) %>% 
  select(1:5)

```

#### Center_scale, don't change weightings column
#### Weightings column is meant to weight the responses to US demographics.
```{r}
data_4_scaled = data_3_impute %>% 
  recipe() %>% 
  update_role(everything(),
              new_role = "predictor") %>% 
  update_role(weight, new_role = "weights") %>% 
  step_center(all_predictors()) %>% 
  step_scale(all_predictors()) %>% 
  prep(training = data_3_impute) %>% 
  bake(new_data = data_3_impute)
```


# 2. Unsupervised clustering with kmeans
#### Prepare data: resample rows by the US demographic weightings.
```{r}
data_4_scaled = data_4_scaled %>% 
  mutate(weight = round(weight / min(weight)))

sample_freqs = data_4_scaled$weight

# drop the weight column before resampling to reduce memory
data_4_scaled = data_4_scaled %>% 
  select(-weight)

sample_index = 1:length(sample_freqs) %>% 
  map(function(i){
    rep(i, sample_freqs[i])
  }) %>% 
  unlist()

data_5_for_clust = data_4_scaled %>% 
  slice(sample_index)
```

#### Kmeans clustering
#### Determine best number of clusters 
https://uc-r.github.io/kmeans_clustering

```{r}
wss = function(df, k) {
  kmeans(df, k, nstart = 25)$tot.withinss
}

num_clusters = 1:15

set.seed(12345)
wss_values = num_clusters %>%
  map_dbl(wss, df = data_5_for_clust)

wss_results = tibble(num_clusters = num_clusters, wss_values = wss_values) %>%
  mutate(wss_lag = lag(wss_values)) %>% 
  mutate(delta_wss = abs(wss_values - wss_lag))

wss_results %>% 
  ggplot(aes(x = num_clusters, y = wss_values)) +
  geom_point() +
  geom_line()

wss_results %>% 
  na.omit() %>% 
  ggplot(aes(x = num_clusters, y = delta_wss)) +
  geom_point() +
  geom_line()
```

#### Cluster data by 5 clusters and visualize
#### Determine distinctive features
https://cran.r-project.org/web/packages/ggfortify/vignettes/plot_pca.html


```{r}
set.seed(12345)
final_kmeans = data_5_for_clust %>%
  kmeans(5, nstart = 25)

data_6_results = data_5_for_clust %>%
  mutate(Cluster = final_kmeans$cluster) %>%
  distinct()

pca_results = prcomp(data_6_results %>% 
                  select(-Cluster),
       center = FALSE, scale. = FALSE)

autoplot(pca_results, 
         data = data_6_results %>% 
           mutate(Cluster = factor(Cluster)), colour = 'Cluster',
         ) +
  ggtitle("Clusters mapped to PC1, PC2")

autoplot(pca_results, 
         data = data_6_results %>% 
           mutate(Cluster = factor(Cluster)), colour = 'Cluster',
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 3
         ) +
  ggtitle("Feature Loadings on PC1, PC2")
```

#### Graphs are a little hard to read but 
#### Trend is observed between low-PC1/hi-PC2 :: hi-PC1/low-PC2

#### Determine the questions driving the features
```{r}
pca_results$rotation %>%
  as_tibble() %>% 
  select(PC1, PC2) %>% 
  mutate(cols = names(data_5_for_clust)) %>% 
  mutate(distance = sqrt(PC1^2 + PC2^2)) %>% 
  arrange(desc(distance))
```
#### So just from an unsupervised perspective, the largest breaks in the survey population (reweighted to US demographics) are race and education level.
#### The biggest respondents group appear to be highly educated, non-minority.


# Supervised learning: Predicting what does masculinity look like?
#### Set target variable: "How masculine are you?"
#### Remove other questions essentially re-asking the same question.
#### Random forest model will provide feature importances.
```{r results=FALSE}
### Remove other questions synonymous with masculinity
data_7_for_regress = data_5_for_clust %>% 
  select(-contains("q0002"),
         -contains("q0004"),
         -contains("q0005"),
         -contains("educ"))

masculin_model = ranger(data = data_7_for_regress
                        , q0001~.
                        #, dependent.variable.name = "q0001"
                        , save.memory = TRUE
                        , importance = 'permutation')

masc_model_importance = importance_pvalues(masculin_model
                                           , method = "altmann"
                                           , data = data_7_for_regress
                                           , formula = q0001~.
                                           , num.permutations = 100
                                           )
```
```{r}
masc_model_importance %>% 
  as.data.frame() %>% 
  mutate(query = row.names(.)) %>% 
  arrange(desc(importance)) %>% 
  slice(1:10)
```

#### Create partial dependency plots for the top 10 variables.
https://bgreenwell.github.io/pdp/articles/pdp.html


```{r message=FALSE}
library(pdp)

top_columns = masc_model_importance %>% 
  as.data.frame() %>% 
  mutate(query = row.names(.)) %>% 
  arrange(desc(importance)) %>% 
  slice(1:10) %>% 
  select(query) %>% 
  unlist()

top_columns_translated = 
  c("7_6. Have relations with women",
    "18. Try to pay when on a date", 
    "7_8. Watch sports", "7_11. Feel lonely",
    "7_7. Have relations with men", 
    "29. Highest grade school completed", 
    "8_3. Your hair or hairline", 
    "7_5. Physically fight someone", 
    "20_4. How gauge interest intimacy? Every situation different.",
    "34. Salary demographic")

wrapper <- function(x, ...) 
{
  paste(strwrap(x, ...), collapse = "\n")
}

pdp_graphs = map(1:length(top_columns),
                 function(pred_var_num){
                   masculin_model %>%
                     pdp::partial(pred.var = top_columns[pred_var_num]) %>%
                     autoplot(rug = TRUE,
                              train = data_7_for_regress) +
                     theme_light() +
                     ggtitle(wrapper(top_columns_translated[pred_var_num],
                                     width = 20))
                   })

grid.arrange(grobs = pdp_graphs[1:6], ncol = 3)
grid.arrange(grobs = pdp_graphs[7:10], ncol = 3)
```


#### Questions, by importance:
#### 7_6. Have relations with women
#### 18. Try to pay when on a date
#### 7_8. Watch sports
#### 7_11. Feel lonely
#### 7_7. Have relations with men
#### 29. Highest grade school completed
#### 8_3. Your hair or hairline
#### 7_5. Physically fight someone
#### 20_4. How do you gauge someone's interest in intimacy? Every situation different.
#### 34. Salary demographic

#### Consider the shapes of the respones. Yes, more masculine men have more sexual relations with women, but compare which questions show a fairly linear response vs. other patterns.
#### For instance, any level of sexual contact with other men results in a complete lack of masculine identity.
#### Surprisingly: A moderate sports watching only results in moderate masculinity. I would have expected masculinity to increase faster with sports watching, similar to the physical fighting response.
#### Masculine men also admit to some level of loneliness.
#### Masculinity seems to require a moderate level of education. However, all but the lowest level of salary earnings feel masculine.